# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

### 1. Load the data (i.e. read.csv())

```{r}
data <- read.csv("activity.csv");
summary(data)
head(data)
```

### 2.Process/transform the data (if necessary) into a format suitable for your analysis

No need to tranform data at this time. NA are ignored.

## What is mean total number of steps taken per day?

Note: For this part of the assignment, the missing values in the dataset are ignored.

To get the total number of steps per day the values of steps have to be summed up (aggregated) up over all intervals and grouped by date.

```{r}
stepsPerDay <- aggregate(steps ~ date, data, sum)  
```

To get the total number of steps per interval the values of steps have to be summed up (aggregated) over all days and grouped by interval.

```{r}
stepsPerInterval <- aggregate(steps ~ interval, data, sum)  
```

### 1. Make a histogram of the total number of steps taken each day

```{r}
hist(stepsPerDay$steps, breaks=20, main="Histogram of Total Steps per Day",xlab="# of Steps", ylab="Frequency")

```

### 2. Calculate and report the mean and median total number of steps taken per day

```{r}
meanStepsPerDay <- mean(stepsPerDay$steps)
medianStepsPerDay <- median(stepsPerDay$steps)
```

Mean and median total number of steps taken per day are:
  
  - mean:   `r meanStepsPerDay`
  
  - median: `r medianStepsPerDay`


## What is the average daily activity pattern?

```{r}
#countPerInterval <- aggregate(steps ~ interval, D, sum)  
avgStepsPerInterval <- aggregate(steps ~ interval, data, mean)  
```

### Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r, echo=FALSE}
plot(avgStepsPerInterval$interval, avgStepsPerInterval$steps, type = "l", main="Avg # of Steps per Interval", xlab="Minute of Day", ylab="# of Steps")
```

### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?


```{r }
idx <- which.max(avgStepsPerInterval$steps)
maxRow = avgStepsPerInterval[idx,]
maxValue <- avgStepsPerInterval$steps[idx]
intervalOfMaxAvgValue <- avgStepsPerInterval$interval[idx]
```

The maximum number of steps of `r maxValue`,  on average across all days, 

occurend on `r intervalOfMaxAvgValue` 


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
